---
title: "Class 10 - In class exercises"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
---
#notes CLASS 10 AND 11

when need to see number of obs in a group

summarise(nObs = n())
nObs is a name we made up for a new column, but n() just counts the number of obs when you have no args
depends on having group_by done first

Note from lecture - for weather
origin month day hour group still does not return unique values
time_hour does better, hour creates daylight_savings duplicates

NOTE - in weathered_flights in lecture, the first plot we get looks weird and maybe sample size is the problem

TIDYVERSE lets you skip quotes around column names. This is "Quasi-quotation"
for selecting in base r, describe the rows (value criteria) you want to select first, comma, then the columns


when you do a left join with an x and a y when the primary key column/s are not unique in terms of row, it creates new rows that are
all combinations of each nonuniquerow on each side of the join. If a key value shows up 2 times in each table, you'll have four rows for that in the new table. If 3 in each side, 9 ultimate rows.


NOTE IN Lecture Hari shows us how to keep outliers by changing the data values such that gusts over 40 are reset to 40. We see that this 
strengths the positive correlation we find because the increase occurs over a smaller x-axis space. Thus Hari doesn't like doing this.
He wants to drop the outlier observations. Wind-gust + dep delay.

Not facet wrapping for origins leads to in your scatterplot multiple y values per x value

##Task 1: Examine the weather data
1. Use the `skim()` function from the `skimr` package to visually display all the variables in weather

```{r}
#install.packages("skimr")
library(nycflights13, tidyverse)
library(skimr)
skim(flights)
```


2. Use `nrow()` to display the number of rows in the weather data

```{r}
nrow(weather)
```


3. Use `dim()` to display the rows and columns of the weather data

```{r}
dim(weather)
```


4. Use `names()` to display the column names of the weather data

```{r}
names(weather)
```


5. Display the `head()` of the weather data
```{r}
head(weather)
```

6. Display the `tail()` of the weather data

```{r}
tail(weather)
```

7. Display a random 10 rows from the weather data using `sample_n()`
```{r}
library(tidyverse)
sample_n(weather, 10)
```


8. What are the primary keys of the weather data? i.e. which column(s) can be used to uniquely identify each observation in the weather data? Hint: `group_by()` with `filter(n() > 1)` to check if there are more than one observation per group.

ME: as is this returns values for these groups, which means they do have N > 1, more than 1 obs per value, so this is not primary key
see below
There is no single var primary key, use origin + time hour
```{r}
weather %>% 
  group_by(time_hour) %>% 
  filter(n() > 1)
```


##Task 2: Explore the indepedent variable
Use a combination of text, tables and charts to explore the nature of the independent variable
ME: should have used skimr! This isn't much info
```{r}
#weather %>% 
 # summary(wind_gust)
summary(weather$wind_gust)

```
```{r}
  ggplot(data = weather, aes(wind_gust)) +
    geom_histogram()
```

```{r}
ggplot(data = weather, aes(wind_gust)) +
  geom_density()
```
Next we sort wind_gist into bins for another way to summarize it.
see the nObs = n trick to count observations
```{r}
weather %>% 
  filter(wind_gust > 0) %>%
  mutate(binned_delay = cut_width(wind_gust, width = 3)) %>% 
           group_by(binned_delay) %>% 
           summarise (nObs = n()) 
         
```


##Task 3. Explore a relationship
Explore the relationship between departure delay and a variable of your choice.

##Task 4: Explore
Explore a multivariate relationship between dep_delay and a combo of vars.
I think the point here is we just take the key and the variable of interest from flights.

```{r}
library(nycflights13) 
library(tidyverse)

WFlll <- flights %>% 
  select(origin, time_hour, dep_delay) %>%
  left_join(x = .,
            y = weather,
            by = c("origin", "time_hour"))

WFlll %>% 
  str()


WFlll %>%  
  mutate(grpdewp = cut_width(dewp, 10), presgrwp = cut_width(pressure, 8)) %>% 
  group_by(grpdewp, presgrwp) %>% 
  summarise (nObs = n(), meandelay = mean(dep_delay, na.rm = T)) %>% 
   ggplot(aes(x = grpdewp, y = meandelay)) +
  geom_point() +
  facet_wrap(~ presgrwp)



  
weather %>% 
  ggplot(aes(x = pressure)) +
  geom_histogram()
           
```

