---
title: "Class 1 Exercises"
output: html_notebook
---

##My first exercise
```{r}
myFirstVar <-  "Funk"
myFirstVar
```

```{r}
#create some atomic vectors
petNames <- c("Tommy", "Fred", "Bobby")
petAge <- c(4, 5, 8)
nale <- c(TRUE, FALSE, FALSE)
```
Determine type of vectors using typeof

```{r}
typeof(petNames)
typeof(petAge)
typeof(nale)
```

```{r}
##create sequence of even numbers from 2 to 40
evenSeq <- seq(from = 2, to = 40, by = 2)
pasteChars <- paste("Hi!")
hiPets <- paste("Hi!")

```

what happens when we combine data types under the combine function?
R expects the atomic vector to be homogenous, so it converts the other values into the appropriate 
type. 

```{r}
c("Hi", 1, "Yo")
```
note the quotes. 1 is converted into string "1"

```{r}
c(T,F,"TRUE")
```
cannot convert string to logical, but logical can convert to strings.
T is short hand for TRUE, which is a built in type "logical"
```{r}
c(T,F,1)
```
you can convert logicals to integers. There's a coercion hierarchy that dictates this.
coercion is the forced conversion of some types to other types.
something about our syntax defines combine function as a vector. For our purposes pretend it only creates 
vectors.

Now lists

```{r}
listA <- list("Thing1", "Thing2", "Thing3")

```
List values can contain anything, including data frames and other lists, nesting is fine

CLASS ASSIGNMENT
1. Create and store a character vector (of your choice), a numeric vector (interger/double) and a factor, all of the same length

```{r}
CVec2 <- c("Hot Dog", "Hamburger", "Ribs")
CVec1 <- c(4L, 6L, 9L)
```


2. Use `seq()` to generate an atomic vector containing all the odd numbers less than 100
```{r}
ClassSeq <- seq(from = 1, to = 99, by = 2)
```

3. Store all the vectors created in step 1 in a named list
```{r}
Sin1List <- list(CVec1, CVec2)
```

4. Create a dataframe using all the vectors stored in step 1
```{r}
ClassDF1 <- data.frame(First = CVec1, second = CVec2)
#can also skip First = part, that declares the column name
#Igot an error when my column name had an integer in it
```

5. Predict the output of the following without running the code in `c("FALSE", FALSE)`, `c(1.0, FALSE)`, `c(2, TRUE)` and `c("two", 2)`

Character
Numeric / Double
Numeric / Integer
Character


6. What get returned when you check the `typeof()` the dataframe created in step 4. Why?

```{r}
typeof(ClassDF1)
```
The answer is that it returns as a list because technically a DF is a type of list.

##Notes
Files tab
Plots Panel

There is a raisehand button in the participants tab in zoom. There's a hover menu on bottom that brings up 
the hand

button next to Rstudio top left makes new notebook. note book has preview button.
Markdown used to create notebooks.
Notebooks can generate HTML pages with no extra effort when written in markdown, weaving in R code.

The links of Hari to a browser, his webpages, there's a button that will let us download the code.
I think you right-click anywhere on the webpage. No, there's a code drop-down button.

shortcuts
Alt+ "-" assignment operator. Neat!
"Chunk" Ctrl +Alt+I. Chunk of code in notebook.
Run a line: Ctrl+Enter (Windows, maybe) where cursor is
ctrl+shift+enter, run chunk

Broom button clears environment
rm(var) 
deletes var

Var <- "var"

R does not pass values. If multiple vars assigned to same value, all don't change when one reassigned.

Three dashes start YAML header at top of notebook. 2nd line can change format
just type the var name to print it again

NOTES
3 main data structures. 1st is vector. think of an array (but arrays are diff)
next is table. see Hari's matrix of data structures in notes.
1d: atomic vector, list vector
2d: Matrix, data frame
Nd:
First entry is homogenous types, latter is heterogeneous
weirdly, lists can have values, each of which can be a vector, each of its own data type.
one data type per value/vector.
Atomic vectors have a direction, i.e. an order. The elements are numbered. Lists have no
direction or intrinsic order.
command. can add "l" maybe to define value as integer?
'c()' is short for combine. combines into a vector
