---
title: "Class 9 - In class exercises"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
---

##Task Set 1: dplyr recap
Time to complete: 8 mins

1. Complete the code below to select the first 3 columns, `dep_delay` and `carrier` columns from flights data and filter all flights that were delayed by more than 2 hours.
```{r, message=FALSE}
library(tidyverse); library(nycflights13)
flights %>% 
    select(1:3, "dep_delay", "carrier") %>% 
    filter(dep_delay > 120, !is.na(dep_delay))
#filtering out NA here but not needed, filter is smart enough
#filter ONLY takes comparisons
  NA>120
  
```

2. Complete the following code to create a new variable that is the difference between `dep_delay` and average monthly delay for each origin airport.
```{r}
flights %>% 
    group_by(origin, month) %>% 
    mutate(MeanDelay =  mean(dep_delay, na.rm = T), DelayDeviation = dep_delay - MeanDelay) %>% 
    select(origin, month, DelayDeviation, dep_delay, MeanDelay)
    
    #after grouping by, operations are performed automatically on aggregates of the values of the 
  #column you have grouped on

```

3. Explain how this code works. Why does it have 3 rows in its output? Answer below the table output.
```{r}
flights %>% 
    group_by(origin) %>% 
    filter(dep_delay == max(dep_delay, na.rm = T)) %>% 
    ungroup()
```

ANSWER: Because there are only 3 origin airports and max returns a single value for each of the values of origin.


##Task Set 2: Keys
Time to complete: 10 mins

Identify the keys in all the datasets included in the planes data (2 of these were discussed in the prep)

1. airlines
2. airports
3. planes
4. weather
5. flights

##Task Set 3: Joins
Time to complete: 15 mins

1. Complete the code below to see if there are any observations in the `flights` data that are not there in `planes`
```{r}
anti_join(flights, planes) %>% 
    distinct(tailnum, carrier)
```

2. Replace the `anti_join()` with a `semi_join()` to see all the flights that have corresponding information in `planes`.

```{r}
semi_join(flights, planes) %>% 
    distinct(tailnum, carrier)
```





3. Complete the code below to find the average departure delay by the year of the planes manufacture. Note that since both the flights and the planes data have a column called year they will be represented as year.x and year.y (x being the first table in the join and y the second). 

```{r}
inner_join(flights, planes) %>% 
    group_by(year) %>% 
    ##only keep flights above a certain threshold
    filter(n() > 30) %>% 
    summarise(meanDelay = mean(dep_delay, na.rm = T)) %>% 
    ggplot(mapping = aes(x = year, y = meanDelay)) +
    geom_point() +
    geom_smooth(method = "lm")
```


##Task Set 4
Time to complete: 20 mins

1. **Hard** As noted in the class prep, there are duplicate keys in the weather data. Remove these (read the two options in the prep) and then use a `left_join()` to join it to flights. Draw a chart of your choice (you can choose to summarise the data if you want) to show the relationship with precipitation (`precip`)

2. **Hard**: Use the `faa` code to combine the `flights` data with `airports` to study the effect of altitude on arrival delay. 


